---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "#>",
  fig.path = "man/figures/README-",
  fig.retina = 3
)
```

# conjointTools

<!-- badges: start -->
[![Lifecycle:
experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://www.tidyverse.org/lifecycle/#experimental)
[![CRAN
status](https://www.r-pkg.org/badges/version/conjointTools)](https://CRAN.R-project.org/package=conjointTools)
<!-- badges: end -->

This package contains tools for designing choice based conjoint survey experiments.

## Installation

The current version is not yet on CRAN, but you can install it from
Github using the **devtools** library:
```{r, eval=FALSE}
devtools::install_github("jhelvy/conjointTools")
```

Load the library with:
```{r}
library(conjointTools)
```

## Make experiment designs

Use the `makeDoe()` function to create a design of experiment 

Generate a full factorial design of experiment with three attributes, each with 2 levels:
```{r}
doe <- makeDoe(levels = c(2, 2, 2))
doe
```

You can also give your variables names based on the attributes. Here is an example of a full factorial design of experiment about apples:
```{r}
doe <- makeDoe(
    levels = c(2, 2, 2),
    varNames = c("price", "type", "freshness")
)
doe
```

The `makeDoe()` function can also generate "D", "A", and "I" optimal fractional factorial designs:
```{r}
doe <- makeDoe(
    levels = c(2, 2, 2),
    varNames = c("price", "type", "freshness"),
    type = "D",
    nTrials = 4
)
doe
```

# Make coded survey designs 

Once a design of experiment has been generated, a survey can be generated by sampling from the design. The design eliminates the possibility of duplicate alternatives appearing in the same choice question.

```{r, echo=FALSE}
doe <- makeDoe(
    levels = c(2, 2, 2),
    varNames = c("price", "type", "freshness")
)
```
```{r}
survey <- makeSurvey(
    doe       = doe,  # Design of experiment
    nResp     = 1000, # Total number of respondents (upper bound)
    nAltsPerQ = 3,    # Number of alternatives per question
    nQPerResp = 6     # Number of questions per respondent
)

dim(survey)
head(survey)
```

The resulting data frame includes the following additional columns:

- `respID`: Identifies each survey respondent.
- `qID`: Identifies the choice question answered by the respondent.
- `altID`:Identifies the alternative in any one choice observation.
- `obsID`: Identifies each unique choice observation across all respondents.

## Assess required sample sizes

To evaluate the necessary sample size to achieve a certain amount of precision about estimated model parameters, use the `sampleSizer()` function with the survey data frame returned from the `makeSurvey()` function. The function estimates a logit model on multiple subsets of the `survey` data frame with increasing sample sizes determined by the `nbreaks` argument:
```{r, fig.width=6, fig.height=4}
results <- sampleSizer(
    survey   = survey,
    parNames = c('price', 'type', 'freshness'),
    parTypes = c('c', 'd', 'd'), # Set continuous vs. discrete variables
    nbreaks  = 10
)

# Preview results
head(results)

# View a plot of the results
sampleSizePlot(results)
```
